The Dynamics of Anthropomorphism in Human-Robot Interaction

### 2014-04-04 paper.tex
discussion section

\subsubsection{Developmental psychology research on the attribution of animacy and intention}
\label{sec:developmental-expl}

@@

\subsubsection{Cognitive science research, the mental model theory (Theory of Mind)
and neural correlates of anthropomorphism}
\label{sec:cognitive-expl}

@@

\subsubsection{Social psychology research: a 3-factor theory of anthropomorphism}
\label{sec:psychological-factors}

@@

The \emph{Stabilized Level of Anthropomorphism} describes the long-term lasting level of
anthropomorphism.

We proposed that \ICA is built on three factors: user's \emph{personality},
robot's \emph{design} and interaction \emph{purpose} (or \emph{interaction
context}). The user's personality and the context of use do also influence \SLA.
In particular, it appears that the user's level of acquaintance with
technologies plays an important role in long-term tendency to
anthropomorphize~\citep{fink_living_2013} (people more familiar with technology
understand, and hence predict, better the behavior of the robot, which in turn
leads them more frequently to ultimately consider the robot as a simple tool).

@@

With this extended understanding of anthropomorphism being determined by multiple factors and showing dynamics over time, it also becomes questionable if the metrics and techniques that are currently used to measure anthropomorphism are sufficient.

Also, the methods and measurements applied to study anthropomorphism appear insufficient, when applying a holistic understanding of anthropomorphism emerging from an interaction between a human user and a robot, for instance. Our point here is that to study a phenomenon as complex as anthropomorphism, we believe that one needs to take into account many factors (see Section \ref{sec:multi-factors}), besides looking at the human-like design of the artifact.
However, most of the
experiments that investigate anthropomorphism in robotics are conducted in
controlled lab-settings during short-term human-robot interactions. This
experimental setting can be critical for studying
anthropomorphism and findings might possibly lead to interpretations that would not be generalizable to natural settings or real interaction scenarios. For instance, some studies are about showing pictures of different robots to human subjects and then asking them how "human-like" they would rate these robots. What points of reference do the participants have in this setting? What can you really tell from a picture? I found that some normal people totally misinterpret the size / height of a robot, when it is not indicated. Also, when showing only pictures, movements, behaviors, materials, and so forth are totally disregarded. However, these aspects are also crucial to assess how "human-like" a robot appears. Last but not least, it has been found that "human-likeness" is a quite diverse and abstract term that includes many things. Instead of asking such a big question, \cite{kahn_jr._robotic_2006} suggest to ask for more concrete constructs that are typical or unique of the concept of "human-likeness", such as being hungry or feeling guilty. For instance, \cite{ruijten_introducing_2014} propose a 25 concrete questions to measure anthropomorphism along a scale that accounts for lower and higher levels of it.
Some studies (e.g. exploring characteristics of the "uncanny valley") also seem to contrast two or
more different anthropomorphic or mechanical systems that might not be
appropriate for comparison given the complex context.

Measuring anthropomorphism is challenging because the phenomenon is partially
unconscious, its mediation is largely subject-dependent\fxnote{Should we put
here different interaction profiles from the Ranger study?}, and we found the
phenomenon generally difficult to verbalize for users.

@@

We present in this section the methodology and techniques we have relied on to
assess anthropomorphic projections in two unrelated studies, the Roomba
experiment~\cite{fink_living_2013}, and the Domino study.

\paragraph{Quantitative assessment of anthropomorphism} Quantitative data are
acquired by systematic annotations of human- robot interactions on video records
of the experiments. Six classes of interactions were identified as reflecting
active commitment of the users in the interaction (gestures, touches, talk to
the robot, show something to the robot, misbehave, look at the experimenter
because worried by the robot's behavior)

\paragraph{Qualitative assessment of anthropomorphism} The main tools for
qualitative measurement of anthropomorphic effect are questionnaires. We have
explored both Likert-scale and open questions. In the Domino study,
questionnaires were analyzed such as points are attributed for key answers,
leading to a qualitative score:

\begin{itemize}
    \item Spontaneously ascribe feelings or states (2 points) {\it Robot is
        hungry, tired}
    \item Ascribe ``own will'' (1--3 points) {\it Robot could leave room, does not
        always obey, could do a b\^{e}tise...}
    \item View robot as friend (1 point)
    \item Ascribe cognitive abilities (0.5--1 point) {\it Robot is able to see,
        able to hear}
    \item Belief robot has feelings (2 points)
    \item Give anthropomorphic reasons for
        \begin{itemize}
            \item not leaving robot alone (1--3 points) {\it Would be sad when
                left alone}
            \item misbehavior (1 point) {\it Robot wants to make a tour, doesn't
                like to play any longer}
        \end{itemize}
\end{itemize}

@@

@@

To remain tractable, most experiments, and even more when they investigate
long-term interactions, are carried with robots exhibiting relatively simple
behaviors. In our model, the maximum level of anthropomorphism \AntMax is
reached at the end of the novelty effect, and then decrease. A natural question
is whether we can design robots such as \AntMax remains high, possibly higher
than $\ant[t_{novelty~peak}]$ (note that to know if sustaining strong
anthropomorphic interaction is desirable is entirely another question).

Intuitively, the behavioral complexity of the robot, and in particular, the ability
for the robot to evolve over time and adapt its behavior to the user as the user
familiarizes itself with the robot (\ie, \emph{co-familiarize}), seems to play a
key role.

...

@@

Another critical point with the use of human-like form in artificial agents are 
ethical aspects. For instance, we need to consider how far we actually want an 
\textit{artificial agent} to resemble a human -- or ourself. Even though we do not 
discuss these aspects in this article, we think the HRI / robotics community needs 
to consider questions of how far an artificial agent should act ``as if it would be 
social'' toward humans. What implications does an anthropomorphic design of robots 
have? What happens as soon as the artificial agent is considered a \textit{moral 
agent} \cite{sullins_when_2006}? Using \textit{socially interactive robots} in 
contexts involving elderly care or rehabilitation for disabled / handicapped 
persons, autistic children is a controversially discussed topic 
\cite{robins_robots_2005}.

>>>>

\subsubsection{Developmental psychology research on the attribution of animacy and intention\\}
\label{sec:developmental-expl}

@@

\subsubsection{Cognitive science research, Theory of Mind, and neural correlates of anthropomorphism\\}
\label{sec:cognitive-expl}

@@

\subsubsection{Social psychology research: a 3-factor theory of anthropomorphism\\}
\label{sec:psychological-factors}

@@

The \emph{Stabilized Level of Anthropomorphism} describes the long-term lasting level of
anthropomorphism.
We proposed that \ICA is built on three factors: user's \emph{personality},
robot's \emph{design} and interaction \emph{purpose} (or \emph{interaction
context}). The user's personality and the context of use do also influence \SLA.
In particular, it appears that the user's level of acquaintance with
technologies plays an important role in long-term tendency to
anthropomorphize~\citep{fink_living_2013} (people more familiar with technology
understand, and hence predict, better the behavior of the robot, which in turn
leads them more frequently to ultimately consider the robot as a simple tool).

@@

%Also, the methods and measurements applied to study anthropomorphism appear insufficient, when applying a holistic understanding of anthropomorphism emerging from an interaction between a human user and a robot, for instance. Our point here is that to study a phenomenon as complex as anthropomorphism, we believe that one needs to take into account many factors (see Section \ref{sec:multi-factors}), besides looking at the human-like design of the artifact.

With this extended understanding of anthropomorphism being determined by multiple factors and showing dynamics over time, it also becomes questionable if the metrics and techniques that are currently used to measure anthropomorphism are sufficient. We would like to discuss two issues: the scenario in which anthropomorphism is measured, and how, with which means it is assessed.

One needs to keep in mind that anthropomorphism arises in an interaction. We believe that as such, anthropomorphism can be better understood when studied and measured in a real, direct human-robot interaction compared to a mediated interaction. In a direct interaction scenario one could also take the human's behavior into account, as it seems to reflect an anthropomorphic interaction \citep{krach_can_2008,hegel_understanding_2008}.
Of course, the scenario and measurements used depend on the research questions, so a direct human-robot interaction might not always be preferred.

However, most of the
experiments that investigate anthropomorphism in HRI are conducted in
controlled lab-settings during a short-term scenario. For instance, pictures or videos of different robots are shown to human subjects who subsequently fill in a questionnaire, to assess how far the different types of robots were perceived human-like.\footnote{It can already be critical to not indicate the height / size of a robot when it is pictured or presented in a video. We cannot assume that without any point of reference, lay people estimate correctly the height of ASIMO or Nao, for instance, they might think these humanoid robots are the same size as a Barbie doll. Also, when showing a picture, the movements, behaviors, materials of a robot, and so forth are totally disregarded. However, these aspects are also crucial features of a robot, impacting how human-like it appears.}
We think that such
experimental settings have limitations for studying a social phenomenon like
anthropomorphism. The critical question is: can anthropomorphism really be measured in a post-interaction closed questionnaire or by rating scales?
Findings might possibly lead to interpretations that would not be generalizable to natural settings or real interaction scenarios (which is always an issue in social science research).

Regarding the second issue, how and with which means anthropomorphism is measured, the central point is that ``human-likeness'' itself is complex and manifold. It as a quite diverse and abstract concept, and instead of asking such a vague question, whether a robot is human-like or not, \cite{kahn_jr._robotic_2006} suggest to ask for more concrete constructs that are typical or unique of the concept of ``human-likeness''.
For instance, \cite{ruijten_introducing_2014} propose a 25-item questions to measure various concrete aspects of human-likeness. Consequently, anthropomorphism can be modeled along a scale that accounts for lower and higher levels of it.

Still, we also feel that measuring anthropomorphism (or social engagement with a robot) is challenging because the phenomenon is partially
unconscious, its mediation is largely subject-dependent\fxnote{Should we put
here different interaction profiles from the Ranger study?}, and we found the
phenomenon generally difficult to verbalize for users.

@@

We present in this section the methodology and techniques that we used to
assess anthropomorphic projections in the child-robot interaction study, called the Domino study (see Section \ref{sec:field-results}, and Figure \ref{fig:ranger}).

\paragraph{Qualitative and quantitative assessment of anthropomorphism in human's behavior\\}

Anthropomorphism can possibly be observed in how humans behave toward a non-human agent (social human-like engagement). Based on video recordings of the Domino experiment, we acquired both quantitative and qualitative behavioral data that reflects anthropomorphism.

Quantitative data are
acquired by systematic annotations of specific types of actions in the video records of the interaction. Six classes of interactions were identified as reflecting
active commitment and/or an anthropomorphic interaction of the user: gestures toward the robot, touching the robot, direct talk to
the robot, showing something to the robot, mistreating the robot (\eg kicking it), and looking at the experimenter because worried by the robot's behavior.

By considering the number of occurrences, the duration, and consequently the frequency of these types of interaction, we can make statements of how far the robot is anthropomorphized in an interaction (based on quantitative data).

We assessed anthropomorphic interactions qualitatively by looking at specific instances of human-social cues in the human's behavior, such as: facial expressions, observable emotional reactions (\eg surprised, high pitch voice or shouting to the robot), eye-contact with the robot, turn-taking gestures, greeting (waving) gestures. These are just suggestions of actions that might reflect anthropomorphism. These actions partly overlap with the quantitative assessment, however it is generally advisable to combine the two.

\paragraph{Qualitative assessment of anthropomorphism when verbally referring to the robot (reflection)\\}

The main tools for
qualitative measurement of anthropomorphic effect are questionnaires and semi-structured interviews. We have
explored both rating-scales in a questionnaire, and open questions in an interview (inspired from \citep{kahn_jr._robotic_2006,weiss_i_2009}).
In the Domino study, interviews were transcribed and participant's answers to the key questions that were assessing an anthropomorphic projection onto the robot, were analyzed. For instance, one question to see how far children would ascribe moral standing to the robot was \emph{``If you had this robot at home, and you were to leave for 2 weeks of holidays, would it be ok to leave the robot alone at home?''}. A point for anthropomorphism was counted, when the participant's answer indicated an anthropomorphic view of the robot. This process led to a ``qualitative score'' that reflects how far the robot is anthropomorphized. We used the following constructs to measure how far the robot is perceived as a human-like agent:

\begin{itemize}
    \item ~Spontaneously ascribe inner states, {\it eg. robot is
        hungry, tired}
    \item ~Ascribe ``own will'', {\it eg. robot could leave room by itself, does not
        always obey, could do a b\^{e}tise...}
    \item ~Describe robot as friend
    \item ~Ascribe cognitive abilities, {\it eg. robot is able to see,
        able to hear}
    \item ~Belief robot has feelings, {\it eg. robot can be happy or sad}
    \item ~Give anthropomorphic explanation for
        \begin{itemize}
            \item ~Why robot can (not) leave room alone {\it eg. it would be sad when left alone}
            \item ~Misbehavior {\it eg. robot wants to make a tour, doesn't
                like to play any longer}
        \end{itemize}
\end{itemize}

@@

One main challenge that robots with applications in daily life contexts face is to sustain long-term interaction.

@@

To remain tractable, most experiments, and even more when they investigate
long-term interactions, are carried out with robots exhibiting relatively simple
behaviors. In our model, the maximum level of anthropomorphism \AntMax is
reached at the end of the novelty effect, and then decreases. A natural question
is whether we can design robots such as \AntMax remains high, possibly higher
than $\ant[t_{novelty~peak}]$ (note that to know if sustaining strong
anthropomorphic interaction is desirable is entirely another question).

Intuitively, the behavioral complexity of the robot, and in particular, the ability
for the robot to evolve over time and adapt its behavior to the user as the user
familiarizes his/herself with the robot (\ie, \emph{co-familiarize}), seems to play a
key role.

\fixme{JF: Severin, I am not sure what you want to say here.}

@@

After all discussion, there remains the question how far a sustained anthropomorphic (human social) interaction between a human and a non-human can be favored, regarding ethical aspects. We need to rethink carefully the consequences that emotional bonds between humans and robots can have. They might be beneficial on one hand but can also hold negative implications. How far should a robot's design encourage us to lastingly project feelings or intentions to it? Why should a machine even react to us as if it had emotions? Can this be the reality we want to live in? What happens as soon as the artificial agent is considered a \textit{moral agent} \cite{sullins_when_2006}? Also, the use of \textit{socially interactive robots} in 
contexts involving elderly care or in rehabilitation and with autistic children is a controversially discussed topic 
\cite{robins_robots_2005}. Do we want robots to comfort us; robots that make us eventually even falling in love with them?
The anthropomorphic design of robots is critical. Why do we even need this? The HRI / robotics community needs to discuss these and other questions.
